---
title: "Homework 2 taa2158"
author: "Tanmayi Amanchi"
date: "2024-09-30"
output: github_document
---

```{r}
library(tidyverse)
library (readxl)
library (haven)
library(dplyr)
```

## Problem 1
This problem focuses on NYC Transit data. Import the NYC Transit CSV.
```{r}
nyc_transit = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
nyc_transit = janitor::clean_names(nyc_transit)
```

Select the required columns
```{r}
nyc_transit <- nyc_transit %>%
  mutate(routes_served = apply(select(.,route1:route11), 1, function(x) paste(na.omit(x), collapse = ", "))) %>%
  select(line, station_name, station_latitude, station_longitude, entry, vending, entrance_type, ada, routes_served) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

The variables in the nyc_transit data set are line, station_name, station_longitude, station_latitude, entry, vending, entrance_type, ada, and routes_served. The cleaning steps performed so far include selecting only the relevant columns, creating a new variable routes_served that consolidates information about the routes served at each station, and converting the entry variable from character strings to logical values. The resulting data set dimensions are 1,868 × 9. This is considered tidy, as each variable is represented by a column, each observation is a row, and each type of observational unit is a table.

Finding how many distinct stations there are  
```{r}
distinct_stations <- nrow(nyc_transit %>% distinct(line, station_name))
distinct_stations
```
According to this code, there are 465 distinct stations in New York City.

Finding how many are ADA compliant
```{r}
ada_compliant_stations <- sum(nyc_transit$ada == TRUE)
ada_compliant_stations
```
According to this code, there are 468 ada compliant stations.

Proportion of station entrances / exits without vending allow entrance?
```{r}
nyc_transit <- nyc_transit %>%
  mutate(vending = ifelse(vending == "YES", TRUE, FALSE)) 

proportion_without_vending <- nyc_transit %>%
  filter(vending == FALSE) %>%                              
  summarise(proportion = mean(entry, na.rm = TRUE)) %>%   
  pull(proportion)

proportion_without_vending
```
The output of this code gives us 0.3770492. This means that 37.70% of stations/entrances don't have a vending allow entrance.

```{r}
a_train_stations <- nyc_transit %>%
  filter(routes_served == "A") %>%
  distinct(line, station_name)  

num_a_train_stations <- nrow(a_train_stations)

num_ada_compliant_a_train_stations <- sum(a_train_stations$ada == TRUE, na.rm = TRUE)

print(num_a_train_stations)  
print(num_ada_compliant_a_train_stations) 
```
21 distinct stations serve the A train. Of the stations that serve the A train, 0 are ADA compliant.

## Problem 2
Import, read and clean Mr. Trash Wheel data set
```{r}
mr_trash = read_excel("./data/202409 Trash Wheel Collection Data.xlsx", 
                      sheet = "Mr. Trash Wheel",
                      range = "A2:N651") %>%
  janitor::clean_names() %>% 
  mutate(sports_balls = round(sports_balls),
         sports_balls = as.integer(sports_balls),
         type = "mr_trash",
         year = as.numeric(year)
  )
```

Import, read, clean, and organize the data for Professor Trash Wheel 
```{r}
professor_trash = read_excel("./data/202409 Trash Wheel Collection Data.xlsx", 
                             sheet = "Professor Trash Wheel",
                             range = "A2:M120") %>% 
  janitor::clean_names() %>% 
  mutate(type = "professor_trash")
```

Import, read, clean, and organize the data for Gwynnda Trash Wheel 
```{r}
gwynnda_trash = read_excel("./data/202409 Trash Wheel Collection Data.xlsx", 
                           sheet = "Gwynnda Trash Wheel",
                           range = "A2:L265") %>% 
  janitor::clean_names() %>% 
  mutate(type = "gwynnda_trash")
```

Combine to produce a single tidy data set
```{r}
trash_combine = bind_rows(mr_trash, professor_trash, gwynnda_trash) %>% 
  relocate(type)
```

There are `r nrow(trash_combine)` observations and `r ncol(trash_combine)` columns in the resulting data set. By using the print function, we can see the first couple rows of the combined data set:

```{r}
print(trash_combine)
```

Finding the total weight of trash collected by Professor Trash Wheel and the total number of cigarette butts collected by Gwynnda in June of 2022
```{r}
prof_weight = 
  trash_combine %>% 
  filter(type == "professor_trash") %>% 
  pull(weight_tons) %>% 
  sum()

gwyn_cigarette = 
  trash_combine %>% 
  filter(type == "gwynnda_trash",
         year == 2022,
         month == "June") %>% 
  pull(cigarette_butts) %>% 
  sum()
```

The total weight of trash collected by Professor Trash Wheel is `r prof_weight` tons. The total number of cigarette butts collected by Gwynnda in June of 2022 is `r format(gwyn_cigarette, scientific=FALSE)`.






